Static contract checking for Haskell

### 2011-11-09 paper.tex
Paper: clarified the incompleteness of the recursion translation.
The reduction translation is incomplete for proofs that require more than one
unfolding of the recursive definition.  I added an example of this, and
an example showing that the obvious "fix" is unsound.

Where $M - f $ means the content of the module $M$ without $f$'s
definition and $f$'s contract. XXX, TODO Stress that it's not always enough
and that we may have to unroll several times!\note[nc]{??? Why? Need an example here.}

\subsection{Module checking}
A module is a collection of function definitions, data definitions and
contracts. What we want to do is to check that functions satisfy their
contract(s).

>>>>

Where $M - f $ means the content of the module $M$ without $f$'s
definition and $f$'s contract. However, this translation is not
complete: some proofs need to unroll the definition more than
once!\note[nc]{For a simple example, consider a contract that is
  proved by a finite number of reductions, but not by induction.  E.g.
}
  \begin{verbatim}
  isZero y = case y of Z -> True | S _ -> False
  eqNat x y = case x of
    Z -> isZero y
    S x' -> eqNatAux x' y
  eqNatAux x y = case y of
    Z -> False
    S y' -> eqNat x y'

  f x = case x of Z -> True | S x' -> f x'
  f ::: (CF&&{x: eqNat x (S Z)}) -> {y:y} 
  \end{verbatim}
\note[nc]{
  Now, the correct way to prove $f$'s contract is two unfoldings and
  reduction: from $eqNat ~x ~(S ~Z)$ we get $x = S ~Z$, and $f~(S~Z) =
  f~Z = \true$.  However, since we translate $f$ to
}
  \begin{verbatim}
  f x = case x of Z -> True | S x' -> f_rec x'
  \end{verbatim}
\note[nc]{
  we get $f~(S~Z) = \rec{f}~Z$ and we're stuck.

  That example is very contrived, in that the much strong contract $f
  ::: \cfc \to \{y:y\}$ is provable by induction in our translation,
  but the same thing can happen in more realistic examples.  For
  example, if you try to prove $add ::: x:(\cfc \land \{x:gt ~x ~Zero\})
  \to y:\cfc \to \{z:gt ~z ~y\}$, then you eventually get to $gt
  ~(S~(\rec{add} ~Z ~y)) ~y$ and then $gt ~(\rec{add} ~Z ~y') ~y'$, for
  $y = S~y'$. But now you're stuck, because there is no reduction rule
  for $\rec{add}$ allowing you to conclude $\rec{add} ~Z ~y' = y'$.

  OK, so we might hope to simply fully define $\rec{add}$, so that
  computations with $\rec{add}$ can be reduced.  That approach is
  unsound.  It's obvious, because as part of the translation of $f :::
  c$ we get $\rec{f} ::: c$ as an axiom, and so if $f ::: c$ doesn't
  actually hold, then $\rec{f} ::: c$ is contradictory.  For
  concreteness, here's simple example. Consider $f$ as above, but
  change the contract to $c := \cfc \to \{y: not ~y\}$.  Then $f$
  doesn't satisfy $c$, and a context that assumes $\rec{f}$ satisfies
  $c$ is contradictory when $\rec{f}$ is fully defined, by
  e.g. reducing $\rec{f}~Z = \true$.
 }
\subsection{Module checking}
A module is a collection of function definitions, data definitions and
contracts. What we want to do is to check that functions satisfy their
contract(s).
